---
title: "Forecasting 3"
author: "simplymathematics"
date: "February 24, 2019"
output: html_document
---

```{r}
library(fpp2)
library(seasonal)
```


## Problem 6.2
### a
```{r, echo = FALSE}
?plastics
autoplot(plastics)
```
Plastic sales tend to peak in the middle of the month.
### b

```{r, echo = FALSE}
plastics_decomposed <- decompose(plastics, type = 'multiplicative')
autoplot(plastics_decomposed)
```
### c
The decomposition shows the same monthly trend as well as a general upward tick in production (excluding the last month). 
### d
```{r, echo = FALSE}
autoplot(plastics, series = 'Plastics Production') + 
  ggtitle("Sales of plastic (thousands)") +
  autolayer(trendcycle(plastics_decomposed), series = 'Trend') + 
  autolayer(seasadj(plastics_decomposed), series = 'Seasonally Adjusted') + 
  xlab("Year") +
  ylab("Monthly Sales")

```
### e
```{r, echo = FALSE}
new <- plastics
new[1] <- 2000
d.new <- decompose(new, type = 'multiplicative')

autoplot(new, series = 'Plastics Production') + 
  ggtitle("Sales of plastic (thousands)") +
  autolayer(trendcycle(d.new), series = 'Trend') + 
  autolayer(seasadj(d.new), series = 'Seasonally Adjusted') + 
  xlab("Year") +
  ylab("Monthly Sales")
```
The outlier effects the seasonally adjusted value greatly, but does not greatly effect the trend line.
### f

```{r, echo = FALSE}
new <- plastics
new[10] <- 2000
d.new <- decompose(new, type = 'multiplicative')

autoplot(new, sies = 'Plastics Production') + 
  ggtitle("Sales of plastic (thousands)") +
  autolayer(trendcycle(d.new), series = 'Trend') + 
  autolayer(seasadj(d.new), series = 'Seasonally Adjusted') + 
  xlab("Year") +
  ylab("Monthly Sales")
```
The later in the data the outlier is, the less it effects the trend line.
## Problem 6.3
### Load Data
```{r, echo=FALSE}
curl :: curl_download("https://otexts.com/fpp2/extrafiles/retail.xlsx", "retail.xlsx")
retail <- readxl :: read_excel("retail.xlsx", skip = 1)
ts <- ts(retail[,"A3349872X"], frequency = 12, start = c(1982,4))
autoplot(ts)
```

### Decompose the series using X11
```{r, echo = FALSE}
ts.x11 <- seas(ts, x11 = "")
autoplot(ts.x11)
```

As expected, the biggest outlier to air traffic was 2001. Additionally, there has been recent volatility perhaps due to fluctuating oil prices and other travel-related concerns.
